### 2010-02-09 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_mediation.tex
Corrected based on review of DERI contribution
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@528 d8b0c59b-6544-0410-94d9-83a5658a058e

\section*{Executive Summary}
\doublespacing
Ontology mediation is a critical problem in heterogeneous environments, since inter-operating services and agents often use different
ontologies, and the differences between them need to be determined and overcome to allow for a seamless data exchange between these parties.
This is a difficult ask, with which web-based agents must cope.

@@

\begin{tabular}
    {| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
    \hline
    \textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2009} & {\textbf{Submitted}} & {27/02/2009}\\ \hline
    \textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
    \textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
    \textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.2 is the second iteration of the FAST ontology matching approach. The firsty iteration contained a general overview of the literature regarding ontology mediation, paving the way for futher analysis of the scenarios arising in FAST that require ontology mediation. Based on that general overview, this current interation presents the approach to implement ontology matching within FAST. We present the different roles within the platform, tasks and challenges related to them, and the way ontology matching is integrated within each role.}\\ \hline
    \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\
    \hline
    \textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {oszkar.ambrus@deri.org} \\ \cline{4-5}
    \textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495112} \\ \hline
    \textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, ontology mediation, ontology matching, ontology alignment}\\ \hline
\end{tabular}
\end{small}
\newpage

@@

\section{Introduction} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:intro}
\paragraph{Goal and Scope.}
FAST is a visual programming platform allowing business users to build enterprise-class mashups, employing various underlying services and generating new ones. Resources in FAST are described semantically using different ontologies and vocabularies, and can therefore be combined to what we can call ``intelligent gadgets''. In the services used within the FAST environment, the ontologies involved in the creation of gadgets and building blocks come from different parties. These ontologies, though different, will sometimes cover the same domain, making ontology matching necessary, as the means to reconcile differences in the various conceptualisations.

@@

\paragraph{Overall Approach for WPX and relation to other Work Packages and Deliverables.}
Ontology matching will play an important role for the integration of different underlying resources in the user-built gadgets. 
We present our initial research in the implementation of ontology matching in this second iteration of D2.4, after having laid the theoretical ground work of ontology matching in the first iteration. The final version of the deliverable (i.e., D2.4.3) will finalize the implementation of ontology mediation within the FAST platform.

\part{Implementing Ontology Matching within the FAST Platform}
\label{part:d2.4.2}

@@

\part{D2.4.1: Theoretical Foundations for Ontology Mediation}
\label{part:d2.4.1}
\include{fast_d2_4_2_part2}

>>>>

\section*{Executive Summary}
\doublespacing
Ontology mediation is a critical problem in heterogeneous environments, since inter-operating services and agents often use different
ontologies, and the differences between them need to be determined and overcome to allow for a seamless data exchange between these parties.
This is a difficult task, with which web-based agents must cope.

@@

\begin{tabular}
    {| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
    \hline
    \textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2009} & {\textbf{Submitted}} & {27/02/2009}\\ \hline
    \textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
    \textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
    \textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.2 is the second iteration of the FAST ontology matching approach. The first iteration contained a general overview of the literature regarding ontology mediation, paving the way for futher analysis of the scenarios arising in FAST that require ontology mediation. Based on that general overview, this current interation presents the approach to implement ontology matching within FAST. We present the different roles within the platform, tasks and challenges related to them, and the way ontology matching is integrated within each role.}\\ \hline
    \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\
    \hline
    \textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {oszkar.ambrus@deri.org} \\ \cline{4-5}
    \textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495112} \\ \hline
    \textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, ontology mediation, ontology matching, ontology alignment}\\ \hline
\end{tabular}
\end{small}
\newpage

@@

\section{Introduction} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:intro}
\paragraph{Goal and Scope.}
FAST is a visual programming platform allowing business users to build enterprise-class mashups, employing various underlying services and generating new ones. Resources in FAST are described semantically using different ontologies and vocabularies, and can therefore be combined to what we can call ``intelligent gadgets'', that rely on semantically enriched building blocks. In the services used within the FAST environment, the ontologies involved in the creation of gadgets and building blocks come from different parties. These ontologies, though different, will sometimes cover the same domain, making ontology matching necessary, as the means to reconcile differences in the various conceptualisations.

@@

\paragraph{Overall Approach for WP2 and relation to other Work Packages and Deliverables.}
Ontology matching will play an important role for the integration of different underlying resources in the user-built gadgets. 
We present our initial research in the implementation of ontology matching in this second iteration of D2.4, after having laid the theoretical ground work of ontology matching in the first iteration. The final version of the deliverable (i.e., D2.4.3) will finalize the implementation of ontology mediation within the FAST platform.

\pagebreak
\part{Implementing Ontology Matching within the FAST Platform}
\label{part:d2.4.2}

@@

\pagebreak
\part{D2.4.1: Theoretical Foundations for Ontology Mediation}
\label{part:d2.4.1}

\include{fast_d2_4_2_part2}

### 2010-02-09 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_part1.tex
Corrected based on review of DERI contribution
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@528 d8b0c59b-6544-0410-94d9-83a5658a058e

\paragraph{Testing Procedure}
For the Amazon--eBay pair we set up a reference alignment, against which the results are evaluated.
We then ran the matching process for all for methods:
\begin{inparaenum}[(1)]
    \item equality,
    \item Levenshtein distance with a confidence threshold of \texttt{0.33} (meaning that any correspondence having a smaller confidence measure will be excluded),
    \item SMOA distance with a threshold of \texttt{0.5} and
    \item Wordnet distance using a threshold of \texttt{0.5}.
\end{inparaenum}
To apply the results, we rendered an XSLT template to transform an example dataset.

@@

The first row (reference) shows the reference alignment, which, naturally, has both perfect precision and recall. We can observe what intuition has predicted, namely that pure string equality (equality) is far too simple and irrelevant, by only taking identical labels. By using string distances and giving certain thresholds (Levenshtein and SMOA), we can see that the results are much less precise, but have a better recall, since this allows for entities having similar names to be discovered, at the expense of having quite a few incorrect results (lower precision); the thresholds allow for low-scored cases to be eliminated, although this results in the exclusion of some correct correspondences. The last column (JWNL) contains the results of the Wordnet-enabled method, which shows quite an improvement (precision of 0.67 and a recall of 0.75), due to the lexical analysis, which performs a much more relevant comparison of strings, giving a high number of correct results. The precision of the JWNL alignment shows only a tiny drop below the recall value, meaning that the number of incorrect correspondences discovered is small, and the main source of error is from the number of correspondences not discovered.

@@

We can deduce that the results provided are satisfactory, even though the methods used were simple, string-based ones, and the process was completely automated without any user input.
We are therefore confident that through some user assistance or an initial input alignment the tool can achieve 100\% correct results.

>>>>

\paragraph{Testing Procedure}
For the Amazon--eBay pair we set up a reference alignment, against which the results are evaluated.
We then ran the matching process for all for methods:
\begin{inparaenum}[(1)]
    \item equality,
    \item Levenshtein distance with a confidence threshold of \texttt{0.33} (meaning that any correspondence having a smaller confidence measure will be excluded),
    \item SMOA distance with a threshold of \texttt{0.5} and
    \item Wordnet distance using a threshold of \texttt{0.5}\footnote{Thresholds were selected based on suggestions in the Alignment API documentation}.
\end{inparaenum}
To apply the results, we rendered an XSLT template to transform an example dataset.

@@

The first row (reference) shows the reference alignment, which, naturally, has both perfect precision and recall. We can observe what intuition has predicted, namely that pure string equality (equality) is far too simple and irrelevant, by only taking identical labels. By using string distances and giving certain thresholds (Levenshtein and SMOA), we can see that the results are much less precise, but have a better recall, since this allows for entities having similar names to be discovered, at the expense of having quite a few incorrect results (lower precision); the thresholds allow for low-scored cases to be eliminated, although this results in the exclusion of some correct correspondences. The last column (JWNL) contains the results of the Wordnet-enabled method, which shows quite an improvement (precision of 0.67 and a recall of 0.75\footnote{The scores are orientative, giving a general picture of how the process approaches 100\% efficiency}), due to the lexical analysis, which performs a much more relevant comparison of strings, giving a high number of correct results. The precision of the JWNL alignment shows only a tiny drop below the recall value, meaning that the number of incorrect correspondences discovered is small, and the main source of error is from the number of correspondences not discovered.

@@

We can deduce that the results provided are satisfactory, even though the methods used were simple, string-based ones, and the process was completely automated without any user input.
We are therefore confident that through some user assistance or an initial input alignment the tool can achieve 100\% correct results, which is our aim for the tool to be used in practice.
